Estimating real-time high treat footfall from Wi-Fi probe requests

### 2018-06-12 abstract/abstract.tex
reorganised workspace

	In the past decade, Wi-Fi has emerged as the most commonly used technolo
gy in providing internet access to mobile devices such as smartphones, tablets a
nd laptops in public and private spaces. This has resulted in multiple Wi-Fi net
works being available at almost every location in urban environments. Traversing
 through this overlapping mesh of Wi-Fi networks, modern mobile devices with Wi-
Fi antennae regularly broadcast a special type of signal known as probe requests
, in order to discover available Wi-Fi networks and switch seamlessly between th
em. This is a hardware level signal which was standardised by IEEE 802.11b/g spe
cification, to relay information about the source mobile device to any Access Po
int (AP) available around it. Since this is the first step in establishing a con
nection between any two devices, it is a standard feature of any device which us
es a Wi-Fi radio to communicate. This makes these probe requests an open, passiv
e, continuous, and wireless source of data available at any urban location. More
over, the data can provide us with clues to understand the number of people pres
ent in the immediate surrounding in real-time and with high granularity \citep{f
reud2015,konto2017}. In this paper, using a set of probe requests collected at a
 high street location in London along with manually counted data, we demonstrate
 that pedestrian footfall can be estimated with considerable accuracy without in
fringing on the privacy of the mobile users involved.

@@

	There have been numerous attempts at using Wi-Fi to measure the volume a
nd movement of people in the built environment for various applications \citep{z
arim2006,sap2015,reki2007}. Though most research obtains feasible and favorable 
results, in recent years, one of the major challenges faced in such attempts has
 been the MAC address randomisation process. This process aims to protect the us
ers’ privacy by anonymising the only globally identifiable portion of the prob
e requests, which results in a set of probe requests generated by the same devic
e with different random MAC addresses \citep{green2008}. There have been various
 successful attempts by researchers to breaking this randomisation process in or
der to extract real MAC addresses, \citep{martin2017} but this usually results i
n serious risk of infringement of the privacy of the users of the mobile devices
. There is a clear gap in the research for exploring methodologies which enable 
us to estimate the number of unique mobile devices from a set of anonymised prob
e requests, without the need to reveal their original MAC addresses.

>>>>

	In the past decade, Wi-Fi has emerged as the most commonly used technolo
gy in providing internet access to mobile devices such as smartphones, tablets a
nd laptops in public and private spaces. This has resulted in multiple Wi-Fi net
works being available at almost every location in urban environments. Traversing
 through this overlapping mesh of Wi-Fi networks, modern mobile devices with Wi-
Fi antennae regularly broadcast a special type of signal known as probe requests
, in order to discover available Wi-Fi networks and switch seamlessly between th
em. This is a hardware level signal which was standardised by IEEE 802.11b/g spe
cification, to relay information about the source mobile device to any Access Po
int (AP) available around it. Since this is the first step in establishing a con
nection between any two devices, it is a standard feature of any device which us
es a Wi-Fi radio to communicate. This makes these probe requests an open, passiv
e, continuous, and wireless source of data available at any urban location. More
over, the data can provide us with clues to understand the number of people pres
ent in the immediate surrounding in real-time and with high granularity \citep{freud2015,konto2017}.
In this paper, using a set of probe requests collected at a
 high street location in London along with manually counted data, we demonstrate
 that pedestrian footfall can be estimated with considerable accuracy without in
fringing on the privacy of the mobile users involved.

@@

	There have been numerous attempts at using Wi-Fi to measure the volume a
nd movement of people in the built environment for various applications \citep{zarim2006,sap2015,reki2007}.
Though most research obtains feasible and favorable 
results, in recent years, one of the major challenges faced in such attempts has
 been the MAC address randomisation process. This process aims to protect the us
ers’ privacy by anonymising the only globally identifiable portion of the prob
e requests, which results in a set of probe requests generated by the same devic
e with different random MAC addresses \citep{green2008}. There have been various
 successful attempts by researchers to breaking this randomisation process in or
der to extract real MAC addresses, \citep{martin2017} but this usually results i
n serious risk of infringement of the privacy of the users of the mobile devices
. There is a clear gap in the research for exploring methodologies which enable 
us to estimate the number of unique mobile devices from a set of anonymised prob
e requests, without the need to reveal their original MAC addresses.
